Efficient ancestry and mutation simulation with msprime 1.0

### 2021-11-17 response-to-reviewers.tex

\begin{point}
Figure 4
It is not clear if these two plots present all the data used for the quadratic
fits, or if these data continue outside of the plots. Samples are described,
but not their ploidy, are they haploids or diploid? From the plots here, it is
not clear if the quadratic relationship holds for sample sizes << 1000. I
suggest this figure should also include a smaller sample size, (eg n=10)
\end{point}

\begin{reply}
TODO
\end{reply}

\begin{point}
Figure 5
I suggest including the absolute Ne and gc rates be included in the legend, in
addition to the scaled rate.
\end{point}

\begin{reply}
We avoided included the absolute values here in order to make the comparison
between the simulators as simple as possible (since SimBac and FastSimBac
required scaled inputs). Since the values we use aren't precisely the estimates
for E.~coli (we rounded some values for simplicity), it would probably
be misleading to include some Ne and gc rates that are consistent with the
scaled value we used, but not equal to the actual E.~coli estimates, and
we would therefore prefer not to include them.
\end{reply}

@@

\end{point}

\begin{reply}
TODO
\end{reply}

\begin{point}
Line 370 - population size(s)
\end{point}

\begin{reply}
Thank you, we have corrected this sentence in the revised manuscript.
\end{reply}

@@

\begin{point}
Occasionally the paper reads more like documentation than a paper, such as when
the Demography class is introduced and several associated functions are
mentioned - this is not relevant for a paper in my opinion.

\end{point}
\begin{reply}
TODO.
\end{reply}

\begin{point}
The section about the development model is interesting - very few papers
include such a section, and indeed I think this is one of the stand-out
features of this project that it is so successfully supported by many
contributors. My only suggestion here is to remove the listing of number of
lines of code - this is a very questionable proxy for quality or quantity of
work, and the work is impressive as it is.

\end{point}
\begin{reply}
Thank you for your suggestion, we have removed the number of lines of code in the revised manuscript.
\end{reply}

### 2021-11-17 response-to-reviewers.tex

\begin{point}
Figure 4
It is not clear if these two plots present all the data used for the quadratic
fits, or if these data continue outside of the plots. Samples are described,
but not their ploidy, are they haploids or diploid? From the plots here, it is
not clear if the quadratic relationship holds for sample sizes << 1000. I
suggest this figure should also include a smaller sample size, (eg n=10)
\end{point}

\begin{reply}
TODO
\end{reply}

\begin{point}
Figure 5
I suggest including the absolute Ne and gc rates be included in the legend, in
addition to the scaled rate.
\end{point}

\begin{reply}
We avoided included the absolute values here in order to make the comparison
between the simulators as simple as possible (since SimBac and FastSimBac
require scaled inputs). Since the values we use aren't precisely the estimates
for E.~coli (we rounded some values for simplicity), it would probably
be misleading to include some Ne and gc rates that are consistent with the
scaled value we used, but not equal to the actual E.~coli estimates, and
we would therefore prefer not to include them.
\end{reply}

@@

\begin{point}
Figure 7
Why is an order of magnitude lower recombination rate used here than in other
examples? This choice seems arbitrary and makes it difficult to relate these
benchmarks results to any others in the paper. Samples are diploid or haploid?
I would suggest the absolute Ne and s rates be included in the legend, in
addition to the scaled rate.
\end{point}

\begin{reply}
TODO Issue 214
\end{reply}

\begin{point}
Figure 8
Averaged points are referenced but not shown "Each point...". The plot could be
more clear where the data points are and where the lines are interpolating
between data points. Because of how the simulations were described and plotted,
I found it awkward to try to compare the running time of the DTWF and
coalescent models, even if this was not the purpose of the current figure. The
figure says "to ensure we are measuring" I would suggest changing the text to
"with a goal of measuring".
\end{point}

\begin{reply}
We have update this figure (and all others in the main text) to include
data markers. We have also rephrased the caption to try to clarify the
parameters simulated and make them consistent with other figures.
\end{reply}

### 2021-11-17 paper.tex
Outline for response to reviewers
Update figures and DTWF caption

\begin{figure}
\begin{center}
\includegraphics{illustrations/example_tree_sequence}
\end{center}
\caption{\label{fig-ts-example} An example tree sequence describing
genealogies and sequence variation for four samples at ten sites
on a chromosome of twenty bases long.
Information is stored in a set of tables
(the tables shown here include only essential columns, and
much more information can be associated with the various entities).
The node table
stores information about sampled and ancestral genomes. The
edge table describes how these genomes are related along a
chromosome, and defines the genealogical tree at each position.
The site and mutation tables together describe sequence variation
% JK: also amongst the ancestral genomes, but only in positions where
% they inherit from the sample. I thought it would be less confusing
% to leave this out?
among the samples.
}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics{figures/sweeps-perf}
\end{center}
\caption{\label{fig-selection-perf}Comparison of selective sweep simulation
performance in \msprime\ and \discoal\ (Intel Xeon E5-2680 CPU).
We report the total CPU time and maximum memory usage when simulating 100
replicates for 10 samples in a model
with a single selective sweep in its history where the beneficial
allele had a scaled selection coefficient of $2Ns=1000$,
a per-base recombination rate of $10^{-9}$ and sequence
length varying from 1kb--200kb.}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics{figures/dtwf-perf}
\end{center}
\caption{\label{fig-dtwf-perf} Comparison of Discrete Time Wright-Fisher
(DTWF) simulation performance in \msprime\ and \ARGON\ (Intel Xeon E5-2680 CPU).
We simulated ancestry for a sample of 1000 haploids from a population of 10000,
and report the (A) total CPU time and (B) maximum memory usage for varying
sequence lengths, and a per-base recombination rate of $10^{-8}$. Each
point is the average over $5$ replicate simulations.
We show observations for \ARGON, \msprime's DTWF implementation (``DTWF'')
and a hybrid simulation of 100 generations of the DTWF followed by
the standard coalescent with recombination model (``DTWF + Hudson'').
Memory usage for \msprime's DTWF and hybrid simulations are very similar.
We ran \ARGON\ with a mutation rate of $0$ and with minimum output options,
to ensure we are measuring only ancestry simulation time.}
\end{figure}

>>>>

\begin{figure}
\begin{center}
\includegraphics{illustrations/example_tree_sequence}
\end{center}
\caption{\label{fig-ts-example} An example tree sequence describing
genealogies and sequence variation for four samples at ten sites
on a chromosome of twenty bases long.
Information is stored in a set of tables
(the tables shown here include only essential columns, and
much more information can be associated with the various entities).
The node table
stores information about sampled and ancestral genomes. The
edge table describes how these genomes are related along a
chromosome, and defines the genealogical tree at each position.
The site and mutation tables together describe sequence variation
among the samples. The genotype matrix and tree topologies shown
on the left are derived from these tables.
}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics{figures/sweeps-perf}
\end{center}
\caption{\label{fig-selection-perf}Comparison of selective sweep simulation
performance in \msprime\ and \discoal\ (Intel Xeon E5-2680 CPU).
We report the average CPU time and maximum memory usage when simulating 3
replicates for 200 samples in a model
with a single selective sweep in its history where the beneficial
allele had a selection coefficient of $s=0.05$,
a per-base recombination rate of $10^{-8}$, population size of $N=10^4$,
and sequence length varying from 100kb--3000kb.}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics{figures/dtwf-perf}
\end{center}
\caption{\label{fig-dtwf-perf} Comparison of Discrete Time Wright-Fisher
(DTWF) simulation performance in \msprime\ and \ARGON\ (Intel Xeon E5-2680 CPU).
We ran simulations with a population size of $10^4$ and
recombination rate of $10^{-8}$, with 500 diploid samples
and varying sequence length.
We report (A) total CPU time and (B) maximum memory usage;
each point is the average over $5$ replicate simulations.
We show observations for \ARGON, \msprime's DTWF implementation (``DTWF'')
and a hybrid simulation of 100 generations of the DTWF followed by
the standard coalescent with recombination (``DTWF + Hudson'').
We ran \ARGON\ with a mutation rate of $0$ and with minimum output options,
with a goal of measuring only ancestry simulation time.
Memory usage for \msprime's DTWF and hybrid simulations are very similar.
}
\end{figure}
